---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# wavcoreinf

<!-- badges: start -->
<!-- badges: end -->

wavcoreinf provides a flexible estimation and evaluation of core inflation from wavelet-based
signal estimation methods. Functionals are developed to facilitate the estimation of several
specifications permitting the selection of the best one according to specific criteria established in
the core inflation literature.

## Installation

You can install  the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("nelson16silva/wavcoreinf", build_vignettes = TRUE)
```
## Example

This is a basic example which shows you how to obtain a good core inflation measure from wavelet. The wavcoreinf's vignette provides more information:

```{r, eval = FALSE}
vignette("wavcoreinf")
browseVignettes("wavcoreinf")
```



```{r example, message = FALSE, warning = FALSE}
library(wavcoreinf)
library(dplyr)
library(tidyr)
library(ggplot2)
library(tibble)
library(purrr)
library(rlang)
library(lubridate)
library(forecast)
library(ellipsis)
```


### Brazilian Headline Inflation and Core Inflation

```{r  Brazilian Headline Inflation and Core Inflation}
n <- c(433, 11427, 16121, 27838, 27839, 11426, 4466, 16122)
x <- c("ipca", "ipcams", "ipcama", "ipcaex0", "ipcaex1", "ipcadp", "ipcaex2", "ipcaex3")
# inf <- getsgs(x, n)
```

### Creating individual time series from 2006-07-01 to 2019-08-01

```{r Creating individual time series from 2006-07-01}
coreinf_br2 <- filter(coreinf_br, date >= "2006-07-01")
date <- coreinf_br2[["date"]]
ts_start <- c(year(date[1]), month(date[1]))
for(i in seq_along(x)) {
  assign(x[[i]], ts(coreinf_br2[[i + 1]], start = ts_start, freq = 12))
}
```

###  Wavelet Core Inflation

```{r Wavelet Core Inflation, message = FALSE, warning = FALSE}
pq <- lags(2, 1)
ebthr_xform <- list(wt = c("dwt", "modwt"))

ebthr_wt <- list(
  wf = c("haar", "la8"),
  n.levels = 3:4
)

ebthr_ebwav <- list(
  vscale = c("level", "independet"),
  a = 3.83,
  prior = c("laplace", "cauchy")
)

ebthr_args <- wav_args_ebthr(ebthr_xform, ebthr_wt, ebthr_ebwav)
ebthr_smooth <- wav_smooth(ipca, ebthr_args)

ebthr_table <- wcore_table(ebthr_smooth, ipca)
ebthr_error_mean <- error_wave_summary(3, x = ebthr_args, y = ipca, lags = pq, k = 15, RMSE = TRUE)

ebthr_summary <- wcore_summary_fcast(ipca, ebthr_table, ebthr_error_mean)
ebthr_best <-  ebthr_summary %>% 
  wcore_summary_best(d = c(1, 1, 1, 1, 2))

ebthr_best_median <- wcore_best_median(ebthr_best)

bayes2_ts <- ts(ebthr_best_median, start = ts_start, freq = 12)
```

### Plot

```{r plot, message = FALSE, warning = FALSE, fig.cap = " Headline and Core Inflation Measures"}
inf <- tibble(
  date = time(ipca), ipca,
  `IPCA-MS` = ipcams,
  `IPCA-EX0` = ipcaex0,
  `IPCA-EX1` = ipcaex1,
  `IPCA-DP` = ipcadp,
  `IPCA-EX2` = ipcaex2,
  `IPCA-EX3` = ipcaex3,
  `IPCA-BAYES2` = bayes2_ts[, 2]
)

inf %>%
  gather(-1:-2, key = "measure", value = "core") %>%
  ggplot(aes(x = date)) +
  geom_line(aes(y = core, linetype = "Core")) +
  facet_wrap(~measure, nrow = 4) +
  geom_line(aes(y = ipca, linetype = "Headline")) +
  labs(x = "", y = "%a.m", linetype = "") +
  theme(legend.position = c(.8, .1)) + # legend.position = "bottom") +
  scale_linetype_manual(values = c(Core = "solid", Headline = "dashed"))   

```
